Quantile Regressions via Method of Moments with multiple fixed effects

### 2023-10-31 article/article.qmd
update
First Draft except conclusions

date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 1p
      cite-style: authoryear
      number-depth: 2

@@

# Methodology

@@

Specifically, given an arbitrary vector of empirical influence functions $\lambda_i(\theta)$, the variance-covariance matrix can be estimated as:
$$\hat{V}(\theta) = \frac{1}{N^2} \sum_{i=1}^N \lambda_i(\theta) \lambda_i(\theta)'
$${#eq-eqvcv}

@@

To show the performance of the extended strategy, we implement the results of a small simulation study. We consider a simple model with a single explanatory variable $x$, and a single fixed effect $g$. The data generating process is as follows:

@@

As expected, @tbl-tb1 shows that point estimates for the point estimates for the quantile regressions are identical to the ones reported in @mss2019 (table 6), including analytical standard errors (GLS). With our estimator, however, we are able to also produce both robust and clustered standard errors for location and scale coefficients. Except for few cases, Robust and Clustered standard errors are larger than GLS standard errors, which may be an indication of misspecification of the model. The GLS standard errors we report differ from the ones in MSS, because they use panel standard errors, equivalent to our clustered standard errors, instead of the analytical standard errors we derive.

@@

In @tbl-tb2, we report the results including both individual and year fixed effects. Because Oil prices only vary across years, the variableis excluded from the model specification. Accounting for time fixed effects, does not change the general conclusions one could make based no the results from @tbl-tb1. The two largest differences are that the log of income percapita has a positive effect on Goverment Surpluses, but only for the 25th quantile, because of the largest impact on the Scale component. Similarly, we observe that the income gap now has an impact on Goverment surplus that is always negative, but increasing across quantiles. In both instances, the effects are not statistically significant.

@@

# Appendix {.appendix}{#sec-appendix}

## Model Identification

@@

$$\begin{aligned}
y = x\beta + \nu \\
\nu = \varepsilon \times x\gamma
\end{aligned}
$$

@@

$$q(y|\tau,X) = x(\beta + q(\varepsilon |\tau) \times \gamma)$$

@@

## Estimation of the variance-covariance matrix

@@

### First Moment Condition: Location Model

@@

### Second Moment Condition: Scale model

@@

### Third Moment Condition: Quantile of Standardized Residual

@@

## Influence functions

### Location coefficients

@@

### Scale Coefficients coefficients**

@@

### Quantile of standardized residual

@@

>>>>

date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: review
      model: 1p
      cite-style: authoryear
      number-depth: 2

@@

# Methodology {#sec-method}

@@

Specifically, given an arbitrary vector of empirical influence functions $\lambda_i(\theta)$, the variance-covariance matrix can be estimated as:

$$\hat{V}(\theta) = \frac{1}{N^2} \sum_{i=1}^N \lambda_i(\theta) \lambda_i(\theta)'
$${#eq-eqvcv}

@@

To show the performance of the extended strategy, we implement a small simulation study. We consider a simple model with a single explanatory variable $x$. In contrast with MSS, we consider a two-way fixed effect structure that is more general than the panel structure used in by the authors. For this excercise, we consider the following data generating process:

$$y_i = \alpha_{1i} + \alpha_{2i} + x_i + (1+x_i + \alpha_{1i} + \alpha_{2i} ) \varepsilon_i
$$

where $\alpha_{1i}\sim \chi^2(1)$, $\alpha_{2i} \sim \chi^2(1)$ and $x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))$ with $\chi_i \sim \chi^2(1)$. We only consider the case when the error term $\varepsilon_i$ is assumed to be normally distributed with mean zero and standard deviation 0.5. We assume that there are 50 mutually exclusive fixed effect groups for each set of fixed effects. Observations are assigned to each subgroup randomly, using a uniform distribution between 1 and 50. We consider a sample sizes of 500, 1000, 2000 and 4000 observations, which implies an average of 10, 20, 40 and 80 observations per group. The model is estimated using the location-scale model with heteroskedastic linear errors, and we report the coefficients for the 25th and 75th quantiles. We run this excercise 5000 times. @tbl-sim1 reports the bias, simulated standard error and mean squared error. We also report the results obtained using an adaptation bias-corrected estimator based on the split-panel jackknife estimator proposed by @dhaene2015.[^jk] These results are labeled JKC.

[^jk]: For the implementation, we estimate the model using the full sample, then, randomly assign each observation into one of two groups, and re-estimate the quantile coefficients for each group. The bias-corrected estimator is then obtained as $\hat\beta(\tau)_{jkc}=2*\hat\beta(\tau)_{full}-0.5*\big(\hat\beta(\tau)_{s1}+\hat\beta(\tau)_{s1}\big)$.

Similar to the findings in MSS, we find that the estimator presents a bias when the sample is small ($N=500$), but this bias shrinks as the sample sizes increases. As MSS describes, the bias seems to be proportional to the sample size, or more precisely to $N/N_{g1}$ and $N/N_{g2}$, where $N_{g1}$ and $N_{g2}$ are the average number of observations per group. Interestingly, the bias-corrected estimator presents an almost zero bias, even when the samples are is small (about 10 observations per sub-group). Also replicating the results of MSS, despite bias reduction obtained using the JKC estimator, the standard errors and mean squared error (MSE) are larger than without correction.

::: {#tbl-sim1 layout-nrow=2 layout-ncol=2}

| N = 500      |         Bias | Sim SE       |         MSE  |
| ------------ | :----------: | :----------: | :----------: |
| q25          |              |              |              |
| \ mmqreg       |        0.104 |        0.247 |        0.072 |
| \ jkc          |        0.008 |        0.297 |        0.088 |
| q75          |              |              |              |
| \ mmqreg       |       -0.091 |        0.251 |        0.071 |
| \ jkc          |        0.002 |        0.303 |        0.092 |

| N = 1000     |         Bias | Sim SE       |         MSE  |
| ------------ | :----------: | :----------: | :----------: |
| q25          |              |              |              |
| \ mmqreg       |        0.050 |        0.173 |        0.032 |
| \ jkc          |       -0.001 |        0.191 |        0.036 |
| q75          |              |              |              |
| \ mmqreg       |       -0.044 |        0.176 |        0.033 |
| \ jkc          |        0.006 |        0.195 |        0.038 |

| N = 1000      |         Bias | Sim SE       |         MSE  |
| ------------ | :----------: | :----------: | :----------: |
| q25          |              |              |              |
| \ mmqreg       |        0.025 |        0.122 |        0.016 |
| \ jkc          |        0.003 |        0.129 |        0.017 |
| q75          |              |              |              |
| \ mmqreg       |       -0.023 |        0.124 |        0.016 |
| \ jkc          |        0.001 |        0.131 |        0.017 |

| N = 2000      |         Bias | Sim SE       |         MSE  |
| ------------ | :----------: | :----------: | :----------: |
| q25          |              |              |              |
| \ mmqreg       |        0.012 |        0.087 |        0.008 |
| \ jkc          |        0.000 |        0.089 |        0.008 |
| q75          |              |              |              |
| \ mmqreg       |       -0.011 |        0.087 |        0.008 |
| \ jkc          |        0.000 |        0.089 |        0.008 |

Bias, Simulated Standard error, and Mean Squared Error

:::

As described in section @sec-se, GLS-Standard errors may not be appropriate if heteroskedasticity in the model is misspecified. This could happen if the heteroskedasticity is not linear in parameters, or if there is correlations within groups. To show the performance of the GLS-Standard errors, we consider two additional simulations. To see the impact of heteroskedasticity misspecification, we assume the DGP is given by:

$$y_i = \alpha_{1i} + \alpha_{2i} + x_i + (1+x_i + \sqrt{x_i} +\alpha_{1i} + \alpha_{2i} ) \varepsilon_i
$$

To show the impact of ignoring correlation within groups, we consider the case where the error $\varepsilon$ is correlated within groups defined by the first fixed effect. Specifically:

$$\varepsilon_i =0.5*(\sqrt{0.5} z_i + \sqrt{0.5} w_g)$$

where $z_i~N(0,1)$ and $w_g~N(0,1)$, with $g$ being the $g_{th}$ group defined by the first fixed effect.

For the simulation, we run the excercise 5000 times, using a sample of size $N=2000$. The results from both simulations are reported in the @tbl-sim2. We report the average coefficient, simulation based Standard errors, and the standard errors using GLS, robust and clustered estimations.

|              |      Het:q25 |      Het:q75 |    Clust:q25 |    Clust:q75 |
| ------------ | :----------: | :----------: | :----------: | :----------: |
| Coefficient  |        0.550 |        1.442 |        0.775 |        1.229 |
| SIM\_SE       |        0.154 |        0.155 |        0.109 |        0.107 |
| GLS\_SE       |        0.476 |        0.451 |        0.288 |        0.221 |
| Robust       |        0.140 |        0.139 |        0.087 |        0.088 |
| Cluster      |         |         |        0.097 |        0.097 |
: Heteroskedasticity misspecification and clustered errors {#tbl-sim2}

One of the conclusions from the results in @tbl-sim2 is that the results for the 25th and 75th quantiles are mostly identical, because of the symmetric error we used in the simulation. On the one hand, because the error structure is misspecified, we should expect to see a large bias in the estimated coefficients, especially in cases of misspecified heteroskedasticity. However, the purpose of this exercise is not to demonstrate the unbiasedness of the coefficients, but rather to assess the performance of the standard error estimation.

As expected, the GLS-standard errors are biased, and the bias is larger in the case of misspecified heteroskedasticity. In fact, with heteroskedastic misspecification, the average GLS-SE (.476) is almost three times as large as the simulation-based standard error (0.154). In this first case, robust standard errors perform better, but with a small difference (0.14 to 0.154). In the case of clustered correlation, GLS-SE also performs poorly. Robust standard errors perform better, but seem to underestimate the magnitude of the true standard errors. Finally, cluster standard errors perform the best, with a small difference between the simulation-based standard errors (0.097 to 0.109).

@@

As expected, @tbl-tb1 shows that point estimates for the point estimates are identical to the ones reported in @mss2019 (Table 6), including analytical standard errors (GLS). With our estimator, however, we are able to also produce both robust and clustered standard errors for location and scale coefficients. Except for a few cases, robust and clustered standard errors are larger than GLS standard errors, which may be an indication of misspecification of the model. The GLS standard errors we report differ from the ones in MSS because they use panel standard errors, which are equivalent to our clustered standard errors, instead of the analytical standard errors we derive.

@@

In @tbl-tb2, we report the results including both individual and year fixed effects. Because oil prices only vary across years, the variable is excluded from the model specification. Accounting for time fixed effects does not change the general conclusions one could make based on the results from @tbl-tb1. The two largest differences are that the log of income per capita has a positive effect on Government Surpluses, but only for the 25th quantile, because of the largest impact on the Scale component. Similarly, we observe that the income gap now has an impact on Government surplus that is always negative, but increasing across quantiles. In both instances, the effects are not statistically significant.

@@

# Appendix {#sec-appendix}

## Derivation of the influence functions

### Model Identification

@@

$$\begin{aligned}
y_i = x_i'\beta + \nu_i \\
\nu_i = \varepsilon_i \times x_i'\gamma
\end{aligned}
$$

@@

$$q(y|\tau,x) = x'(\beta + q(\varepsilon |\tau) \times \gamma)$$

@@

### Estimation of the variance-covariance matrix

@@

#### First Moment Condition: Location Model

@@

#### Second Moment Condition: Scale model

@@

#### Third Moment Condition: Quantile of Standardized Residual

@@

### Influence functions

#### Location coefficients

@@

#### Scale Coefficients

@@

#### Quantile of standardized residual

@@

## Implementation

The method described here can be implemented using any of the following packages

- `mmqreg` in Stata: Installed from Github
- `mmqreg` in R: Programs Available in Github
- `mmqreg` in Python: 
- `mmqreg` in julia:
